---
title: "ML_linear_regression"
author: "Tian Zhang"
date: "12/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r initialize}
library(ggplot2)
library(ggthemes)
library(data.table)
library(dplyr)
```

## Data

The data has the following features:
datetime - hourly date + timestamp
season - 1 = spring, 2 = summer, 3 = fall, 4 = winter
holiday - whether the day is considered a holiday
workingday - whether the day is neither a weekend nor holiday
weather -
        1: Clear, Few clouds, Partly cloudy, Partly cloudy
        2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
        3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
        4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog
temp - temperature in Celsius
atemp - "feels like" temperature in Celsius
humidity - relative humidity
windspeed - wind speed
casual - number of non-registered user rentals initiated
registered - number of registered user rentals initiated
count - number of total rentals

```{r bike}
bike <- fread("bikeshare.csv")
head(bike)
```

## Exploratory Data Analysis

```{r  , echo=FALSE}
count_temp <- ggplot(bike, aes(x = temp, y = count, color = temp)) + geom_point(alpha = 0.5)
count_temp
```

```{r, echo=FALSE}
bike$datetime <- as.POSIXct(bike$datetime)
count_date <- ggplot(bike, aes(x = datetime, y = count)) + 
        geom_point(aes(color = temp),alpha = 0.5) +  scale_colour_gradient(high='red',low = "blue")
count_date
```

```{r correlation}
cor.temp_count <- cor(bike[,c("temp","count")])
cor.temp_count
```

```{r  , echo=FALSE}
count_season <- ggplot(bike, aes(x = season, y = count)) + 
        geom_boxplot(aes(color = factor(season)),alpha = 0.5)
count_season
```

## Feature Engineering
```{r hour}
to_hour <- function(x){
        format(x, "%H")
}
bike$hour <- sapply(bike$datetime, to_hour)
```

```{r workingday hourly}
workingday <- subset(bike, workingday == 1)
count_hour <- ggplot(workingday, aes(x = hour, y = count)) +
        geom_point(aes(color = temp), alpha = 0.5, position = position_jitter(w=1, h=0)) +
        scale_colour_gradientn(colours = c('blue','purple','green','yellow','orange','red'))
count_hour
```

```{r nonworkingday hourly}
nonworkingday <- subset(bike, workingday == 0)
count_hour <- ggplot(nonworkingday, aes(x = hour, y = count)) +
        geom_point(aes(color = temp), alpha = 0.5, position = position_jitter(w=1, h=0)) +
        scale_colour_gradientn(colours = c('blue','purple','green','yellow','orange','red'))
count_hour
```

## Building the Model
```{r temp.model}
temp.model <- lm(count~temp, data = bike)
summary(temp.model)
```

```{r predict}
6.0462 + 9.1705*25
temp.test <- data.frame(temp = c(25))
predict(temp.model, temp.test)
```

```{r hour to number}
bike$hour <- sapply(bike$hour, as.numeric)
```

```{r multi factor regression}
multi.model <- lm(count ~ .-casual-registered-datetime-atemp, bike)
summary(multi.model)
```
